% !TeX root = article-enso.tex

\listfiles
\documentclass[review, 12pt]{elsarticle}

\input{texfiles/setup.tex}

\begin{document}


\begin{frontmatter}

\title{Mechanisms underlying the epipelagic ecosystem response to ENSO in the equatorial Pacific ocean}
%\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}

%% Group authors per affiliation:
\author[mymainaddress]{Nicolas Barrier\corref{mycorrespondingauthor}}
%\address{Radarweg 29, Amsterdam}
\fntext[myfootnote]{MARBEC, University of Montpellier, CNRS, Ifremer, IRD, S\`ete, France}

\author[mymainaddress]{Matthieu Lengaigne}
\author[mymainaddress]{Jonathan Rault}
%\author[jerome]{J\'{e}r\^{o}me Guiet}
\author[renaud]{Renaud Person}
\author[chris]{Christian Eth\'{e}}
\author[renaud]{Olivier Aumont}
\author[mymainaddress]{Olivier Maury}

\cortext[mycorrespondingauthor]{Corresponding author}	

%% or include affiliations in footnotes:
%\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
%\ead[url]{www.elsevier.com}
%\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
%\ead{support@elsevier.com}
\address[mymainaddress]{MARBEC, Univ. Montpellier, CNRS, Ifremer, IRD, Sète, France}
\address[renaud]{LOCEAN, IRD}
\address[chris]{IPSL, CNRS}
%\address[jerome]{Department of Atmospheric and Oceanic Sciences, University of California, Los Angeles}

\begin{abstract}

The \nino{}/Southern Oscillation is known to strongly impact marine ecosystems and fisheries. In particular, \nino{} years are characterized, among other things, by a decrease in tuna catches in the western Pacific and an increase in the central Pacific, whereas these catches accumulate in the far western Pacific during \nina{} conditions. However, the processes driving this zonal shift in the resource (changing habitat conditions, currents or food availability) remain unclear. Here, we use an hindcast simulation from the mechanistic ecosystem model APECOSM that reasonably reproduces the observed zonal shift of the epipelagic community in response to ENSO to understand the mechanisms underlying this shift.


Although the response of modeled epipelagic communities to \nino{} is relatively similar for the different size classes studied, the processes responsible for these changes vary considerably by organism size. One of the major results of our analysis is the critical role of eastward passive transport by \nino{} related surface current anomalies in these changes for all size classes. While the effects of passive transport dominate the effects of growth and predation changes everywhere for large organisms, this is not the case for intermediate-sized organisms in the western Pacific, where the decrease in biomass is first explained by increased predation and then decreased foraging success. For small organisms, changes in growth rate induced by the influence of temperature on fish physiology is an important process that reinforces the biomass increase induced by passive horizontal transport in the eastern Pacific and the biomass decrease induced by increased predation by intermediate organisms near the dateline. Finally, contrary to what is often assumed, our model shows that active habitat-based movements are not required to explain the westward biomass shifts that are observed during ENSO.

This study illustrates the relevance of using a mechanistic ecosystem model to disentangle the role of the different processes controlling biomass changes. It highlights the essential dynamic role of ocean currents in shaping the response of marine communities to climate variability and its interaction with biological (e.g. growth) and ecological (e.g. foraging and predation) processes, whose relative importance varies with organisms' size and contribute to modify the community structure.

\end{abstract}

\begin{keyword}
fish \sep biomass \sep ENSO \sep \nino \sep \nina \sep ecosystem modelling \sep DEB \sep advection \sep growh \sep predation \sep equatorial Pacific \sep APECOSM \sep epipelagic \sep ecosystem \sep habitat
\end{keyword}

\end{frontmatter}

\linenumbers

%\section{Introduction}
%
%Assessing the impacts of \nino\ events on the fish biomass is a valuable way to infer the future state of ocean in warming ocean.\\
%
%\nino\ events are not comparable to each other, they have their own characteristics.
%\newpage

\input{texfiles/intro}
\input{texfiles/data_method}
\input{texfiles/model_val}
\input{texfiles/nino97}
\input{texfiles/decadal}

\section*{Acknowledgement}

The authors acknowledge the Pôle de Calcul et de Données Marines (PCDM, \url{http://www.ifremer.fr/pcdm}) for providing DATARMOR storage, data access, computational resources, visualization and support services.\\
Data analysis and representation has been performed using Python 3.9.4. 
Covariance analysis and detrending has been performed using the \emph{numpy} and \emph{scipy} packages. Plots and maps have been made by using the \emph{matplotlib} and \emph{cartopy} packages. The tuna drawing in Figure \ref{fig:proc_summary} is courtesy of Tchiiweb (http://www.tchiiweb.com/). The Pacific data presented in this study and the scripts used in the analysis are available on Zenodo (\url{https://doi.org/10.5281/zenodo.7142804}).

\listoffigures
\listoftables

\clearpage

\bibliography{ApecosmBib2}

\end{document}
